---
title: "01-download-data"
output: html_document
date: "2023-04-07"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = FALSE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 6,       # Set plot width in inches
  fig.height = 4,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```
This is step 1 of my project workflow. In this step, I am organizing my fastq files and checking their hash values. I originally downloaded my data from a shared google drive folder onto my computer and then uploaded into RStudio. Originally, the fastq files were located in multiple layers of directories.

The URL for step 1 is located here: https://rpubs.com/HailaSchultz/download-data

The URL for step 2 is located here: https://rpubs.com/HailaSchultz/trim-primers


load packages
```{r}
library(knitr)
```


# Move files into one folder

move multiple files from multiple layers of directories into the data directory
```{bash}
cd ../data
find . -name '*.gz' -exec mv {} ../data/fastq-files/ \;
```


create checksums file for all fastq files in data folder
```{bash}
cd /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/data/
pwd
ls
shasum ../data/fastq-files/*.gz > fastq_checksums.sha
cat fastq_checksums.sha
```

check if files match original
```{bash}
cd ../data
shasum -c fastq_checksums.sha
```



