---
title: "04-pair"
output: html_document
date: "2023-05-10"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = FALSE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 6,       # Set plot width in inches
  fig.height = 4,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

# Dereplication
```{r}
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
# Name the derep-class objects by the sample names
names(derepFs) <- sample.names
names(derepRs) <- sample.names
```

# Apply core sample inference algorithm
```{r}
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
```
```{r}
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
```

inspect data-class object
```{r}
dadaFs[[1]]
```

# Merge Paired Reads
```{r}
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
```

# Construct ASV table

make sequence table
```{r}
seqtab <- makeSequenceTable(mergers)
```

inspect sequence length distributions
eventually make histogram here
```{r}
table(nchar(getSequences(seqtab)))
```

# remove chimeras
```{r}
seqtab.nochim <- removeBimeraDenovo(seqtab, method="consensus", multithread=TRUE, verbose=TRUE)
dim(seqtab.nochim)
```
```{r}
sum(seqtab.nochim)/sum(seqtab)
```
chimeras account for very low perentage of reads


# track how many reads made it through the pipeline
```{r}
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(dadaRs, getN), sapply(mergers, getN), rowSums(seqtab.nochim))

colnames(track) <- c("input", "filtered", "denoisedF", "denoisedR", "merged", "nonchim")
rownames(track) <- sample.names
head(track)
```
no step had a majority of the reads removed

