---
title: "05-Blast"
output: html_document
date: "2023-05-11"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = FALSE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 6,       # Set plot width in inches
  fig.height = 4,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

load packages
```{r}
library(dplyr)
library(tidyverse)
```

download software from NCBI
```{bash}
#change directory
cd /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/software
#download software
curl -O https://ftp.ncbi.nlm.nih.gov/blast/executables/blast+/LATEST/https://ftp.ncbi.nlm.nih.gov/blast/executables/blast+/LATEST/ncbi-blast-2.14.0+-x64-macosx.tar.gz
```


unzip
```{bash}
cd /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/software
tar -xzf ncbi-blast-2.14.0+-x64-macosx.tar.gz
```

Check if it's working
I had to go into my computer settings and give permissions to use blast because it is from and "unidentified developer"
```{bash}
/Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/software/ncbi-blast-2.14.0+/bin/blastx -h
```

# Make blast database
Download the reference. The reference used here is from metazoogene and is the file for all marine fauna and flora of the world oceans
```{bash}
#download from url
cd /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/data
curl -O https://www.st.nmfs.noaa.gov/nauplius/media/metazoogene/atlas/data-src/MZGfasta-coi__MZGdbALL__o00__A.fasta
#unzip
gunzip -k MZGfasta-coi__MZGdbALL__o00__A.fasta
```

make the blast database
I changed dbtype from prot to nucl
```{bash}
/Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/software/ncbi-blast-2.14.0+/bin/makeblastdb \
-in /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/data/MZGfasta-coi__MZGdbALL__o00__A.fasta \
-dbtype nucl \
-out /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/blastdb/MZGfasta-coi__MZGdbALL__o00__A
```

```{r}
getwd()
```


Write output to a fasta file that can be blasted
transpose table
```{r}
seqtab.nochim_trans <- as.data.frame(t(seqtab.nochim)) %>% rownames_to_column(var = "sequence") %>% 
    rowid_to_column(var = "OTUNumber") %>% mutate(OTUNumber = sprintf("otu%04d", 
    OTUNumber)) %>% mutate(sequence = str_replace_all(sequence, "(-|\\.)", ""))
```


```{r}
df <- seqtab.nochim_trans
seq_out <- Biostrings::DNAStringSet(df$sequence)

names(seq_out) <- str_c(df$OTUNumber, df$Supergroup, df$Division, df$Class, 
    df$Order, df$Family, df$Genus, df$Species, sep = "|")

Biostrings::writeXStringSet(seq_out, str_c( "Zoop_ASV.fasta"), compress = FALSE, 
    width = 20000)

#I had to move this file from my working directory to my data directory
```


```{bash}
head /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/data/Zoop_ASV.fasta
echo "How many sequences are there?"
grep "TTTATCTGCAGGAATTGCACATGCTGGGGCTTCTGTAGATATAGGAATCTTTTCACTACATATTGCTGGAGCCTCTTCTATTTTAGGAGCAGTTAATTTCATTACAACTGTAATTAATATGCGATCAGCAGGAATAACTATAGACCGTATCCCCTTATTCGTATGGTCTGTGTTTATTACAGCAATTTTACTTTTATTGTCTCTACCAGTTTTAGCTGGAGCAATCACAATACTTTTAACAGACCGTAATTTAAATACATCTTTCTTCGACCCTGCCGGGGGTGGTGACCCTATTTTATATCAACATCTATTC" /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/data/Zoop_ASV.fasta
```

# Run Blast

change blastx to blastn
```{bash}
/Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/software/ncbi-blast-2.14.0+/bin/blastn \
-query /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/data/Zoop_ASV.fasta \
-db /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/blastdb/MZGfasta-coi__MZGdbALL__o00__A \
-out /Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/output/Zoop_ASV.tab \
-num_threads 8 \
-max_target_seqs 1 \
-outfmt 6
```

```{bash}
head -2 ../output/Zoop_ASV.tab
wc -l ../output/Zoop_ASV.tab
```


