---
title: "06-Visualization"
output: html_document
date: "2023-05-12"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,         # Display code chunks
  eval = FALSE,        # Evaluate code chunks
  warning = FALSE,     # Hide warnings
  message = FALSE,     # Hide messages
  fig.width = 6,       # Set plot width in inches
  fig.height = 4,      # Set plot height in inches
  fig.align = "center" # Align plots to the center
)
```

load packages
```{r, eval=TRUE}
library(dplyr)
library(janitor)
library(vegan)
library(data.table)
library(ggplot2)
library(ggpubr)
```


# Visualization


## Filter Blast hits and merge tables

merge Blast IDs with OTU table

```{r, eval=TRUE}
#change directory
setwd("/Users/hailaschultz/GitHub/haila-ZoopMetabarcoding/output")

#read ASV table into R
Zoop_ASV<-read.table("Zoop_ASV.tab")

#rename columns
colnames(Zoop_ASV) = c("OTUNumber", "Species", "pident", "length","mismatch","gapopen","qstart","qend","sstart","send","evalue","bitscore")

#merge tables by otu number
read_table<-left_join(Zoop_ASV, seqtab.nochim_trans, by = join_by("OTUNumber" == "OTUNumber"))
```
Visualize part of the table here


Filter hits
```{r, eval=TRUE}
# remove ASV with sequences shorter than 300 bp
read_table <- subset(read_table, read_table$length>300) 

#remove ASV with pident <95%
read_table <- subset(read_table, read_table$pident>95) 

```

Sum Sequences by Taxa
```{r, eval=TRUE}
#remove unneeded columns
read_table_summed<- read_table[ -c(1,3:13) ]

#remove prefix
read_table_summed$Species <- sub(".*__", "", read_table_summed$Species)

#summarize by species
by_species <- read_table_summed %>%
  group_by(Species)

read_table_summed<-by_species %>%
  summarise_all(sum)
```


## NMDS

Prep table for NMDS
```{r, eval=TRUE}
# transpose
read_table_transposed <- transpose(read_table_summed)
# get row and colnames in order
colnames(read_table_transposed) <- rownames(read_table_summed)
rownames(read_table_transposed) <- colnames(read_table_summed)
#move first row to column names
read_table_transposed<-row_to_names(read_table_transposed, 1, remove_rows_above = FALSE) 
#convert to matrix
community_matrix<-as.matrix(read_table_transposed)
#convert to numeric matrix
community_matrix<- matrix(as.numeric(community_matrix),   
                  ncol = ncol(community_matrix))
#convert to proportions
community_matrix<-community_matrix/rowSums(community_matrix)
#arcsine sqrt transformation
community_matrix<-asin(sqrt(community_matrix))
```

Run NMDS
```{r, eval=TRUE}
#run NMDS
NMDS=metaMDS(community_matrix,distance="bray",trymax=100)
```
```{r, eval=TRUE}
stressplot(NMDS)
plot(NMDS)
NMDS
```
Make envrionmental table
```{r, eval=TRUE}
#export file names
env <- as.data.frame(row.names(read_table_transposed))
#change column name
colnames(env)[colnames(env) == "row.names(read_table_transposed)"] <- "file"

#create year column
env$year <- substr(env$file, 1, 4)
#create month column
env$month <- substr(env$file, 5, 7)
#create station column
env$station <- str_extract(env$file, "P12|P22|P28|P38|P402|P4|P8")
#create run column
env$run <- str_extract(env$file, "48samples|-P1|-P2")
```

extract scores
```{r, eval=TRUE}
data.scores = as.data.frame(scores(NMDS)$sites)
#add environmental columns
data.scores$station = env$station
data.scores$year = env$year
data.scores$month = env$month
data.scores$run = env$run
```

create station plot
```{r, eval=TRUE}
station_plot = ggplot(data = data.scores, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = data.scores, aes(colour = station), size = 3, alpha = 0.5) + 
     scale_colour_manual(values = c("orange", "steelblue","darkgreen","violet","red","darkblue","limegreen")) + 
     theme(axis.title = element_text(size = 10, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 10, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 9, colour = "grey30")) +
     labs(colour = "Station")
     
station_plot

```
There is quite a bit of overlap among stations, but you can see that some stations are distinct from one another. For example, P22 is different from P38, but P38 and P12 are very similar.

create year plot

```{r, eval=TRUE}
year_plot = ggplot(data = data.scores, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = data.scores, aes(colour = year), size = 3, alpha = 0.5) + 
     scale_colour_manual(values = c("orange", "steelblue")) + 
     theme(axis.title = element_text(size = 10, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 10, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 9, colour = "grey30")) +
     labs(colour = "Year")
     
year_plot

```
between these two years, there doesn't seem to be much difference

create month plot
```{r, eval=TRUE}
month_plot = ggplot(data = data.scores, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = data.scores, aes(colour = month), size = 3, alpha = 0.5) + 
     scale_colour_manual(values = c("orange", "steelblue","darkgreen")) + 
     theme(axis.title = element_text(size = 10, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 10, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 9, colour = "grey30")) +
     labs(colour = "Month")
     
month_plot

```
There seem to be some clear seasonal differences along axis 1 - April appears to be distinct

create run plot
```{r, eval=TRUE}
run_plot = ggplot(data = data.scores, aes(x = NMDS1, y = NMDS2)) + 
     geom_point(data = data.scores, aes(colour = run), size = 3, alpha = 0.5) + 
     scale_colour_manual(values = c("orange", "steelblue","darkgreen")) + 
     theme(axis.title = element_text(size = 10, face = "bold", colour = "grey30"), 
     panel.background = element_blank(), panel.border = element_rect(fill = NA, colour = "grey30"), 
     axis.ticks = element_blank(), axis.text = element_blank(), legend.key = element_blank(), 
     legend.title = element_text(size = 10, face = "bold", colour = "grey30"), 
     legend.text = element_text(size = 9, colour = "grey30")) +
     labs(colour = "Run")
     
run_plot

```

## Diversity Indices

```{r}
#convert to matrix
diversity<-as.matrix(read_table_transposed)
#convert to numeric matrix
diversity_matrix<- matrix(as.numeric(diversity),   
                  ncol = ncol(diversity))
#add in row and column labels
colnames(diversity_matrix) <- colnames(diversity)
rownames(diversity_matrix) <- rownames(diversity)
```

get shannon diversity index
```{r}
shannon_scores<-as.data.frame(diversity(diversity_matrix, index="shannon"))
```

add environment columns back in
```{r}
#change column name
colnames(shannon_scores)[1] <- "shannon_index"
#add environmental columns
shannon_scores$station = env$station
shannon_scores$year = env$year
shannon_scores$month = env$month
shannon_scores$run = env$run
```

make violin plot
```{r}
diversity<- ggviolin(shannon_scores, x = "station", y = "shannon_index",
 add = "boxplot", fill = "station") 
diversity
```

Overall, the highest diversity was seen at P22, which is the station located in the Strait of Juan de Fuca. This site is closest to the ocean, so it may contain inland taxa as well as more offshore taxa.


## Taxa plot

get the top 20 most abundant species
```{r}
species_sums <- as.data.frame(colSums(diversity_matrix))
colnames(species_sums)[1] <- "abundance"
species_sums$abundance<-as.numeric(species_sums$abundance)
species_sums <- rownames_to_column(species_sums, "species")
species_sums <-as.data.frame(species_sums[order(-species_sums$abundance),])

#get top 20 species
top_20_species <- head(species_sums, 20)
```

plot
```{r}
ggplot(top_20_species, aes(y=abundance, x=reorder(species, abundance))) + 
    geom_bar(position="dodge", stat="identity")+ theme_bw()+coord_flip()+ylab("total reads")+xlab("species")
```
The samples were dominated by Calanus pacificus (a copepod) and Aegina citrea (a jellyfish). Other abundant species included eutonina indicans and Clytia gregaria, two other jellyfish. Other common copepods like Centropages abdonminalis and Metridia pacifica were also represented
